---
title: "Homework 1 - Big Data in R"
author: "Daniel Simpson"
date: "Sunday, November 24, 2019"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Import the appropriate dataset:
```{r}
library(ISLR)
mydata <- Auto
```
A full summary of the Auto dataset:
```{r}  
summary(Auto)
```

Variance for each varaible in the Auto dataset:
```{r}
sapply(Auto[,-c(7,8,9)], var)
```

Make a plot of all the predictor variables to see any correlations:
```{r echo = FALSE}
plot(Auto[,-c(7,8,9)], lower.panel = NULL)
```

Or we can use a correlation plot to visualize the correlation between each predictor:
```{r}
library(corrplot)
correlation_matrix <- cor(Auto[,-c(7,8,9)])

corrplot(correlation_matrix, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```
This plot allows us to see that the larger the circle, the more of a correlation the predictors have. 
Dark blue indicates a positive correlation, while dark red indicates a negative correlation. 
This allows us to identify easily whether the predictors share a relationship or not. 
Obviously the diagonals can be ignored as they represent the predictors relationship with themselves.



We can further investigate this correlation by ploting the two predictors and assigning a best fit line.
  

```{r echo=FALSE}
lm.fit <- lm(mpg ~ horsepower, data = mydata)
plot(mydata$horsepower, mydata$mpg,
     xlab="horsepower", ylab = "mpg",
     main = "Horsepower vs MPG with regression line",
     ylim = c(0,60), xlim = c(30,280))
abline(lm.fit, col = "red")
legend("topright", legend = "Regression Line", lty = 1, col = "red")
summary(lm.fit)
```
The RSE is equal to 4.906
The R-squared value is 0.6049, so about 60.49% of the MPG can be explained by the regression line.

We can then add confidence and predictive intervals:

```{r}
plot(mydata$horsepower, mydata$mpg,
     xlab="Horsepower", ylab = "MPG",
     main = "Horsepower vs MPG",
     ylim = c(5,50), xlim = c(40,245))
abline(lm.fit, col = "red")
newdata1 <- seq(20, 280, by = 50)
conf_int <- predict(lm.fit, newdata = data.frame(horsepower = newdata1), interval = "confidence", level = 0.99)
pred_int <- predict(lm.fit, newdata = data.frame(horsepower = newdata1), interval = "prediction", level = 0.99)
lines(newdata1, conf_int[,2], col="blue", lty=4)
lines(newdata1, conf_int[,3], col="blue", lty=4)
lines(newdata1, pred_int[,2], col="purple", lty=5)
lines(newdata1, pred_int[,3], col="purple", lty=5)
legend("topright", legend = c("Regression line", "Confidence Interval", "Predictor Interval"), col = c("red", "blue","purple"), lty = c(1,4,5), cex = 0.8)
```

Or we can look at the prediction of MPG when horsepower = 89:

```{r}
predict(lm.fit, data.frame(horsepower = 89))
```

By setting the interval to 99% we can observe the confidence interval at horsepower set to 89:

```{r}
predict(lm.fit, data.frame(horsepower = 89), interval = "confidence", level = 0.99)

```

We can also look at the prediction interval for 99% for horsepower set to 89:

```{r}
predict(lm.fit, data.frame(horsepower = 89), interval = "prediction", level = 0.99)

```


```{r echo = FALSE}
plot(Auto$weight, Auto$mpg, xlab = "Weight", ylab = "MPG", main = "MPG vs Weight")
lm.fit <- lm(Auto$mpg ~ Auto$weight, Auto)
abline(lm.fit, col = 'green')
legend("topright", legend = "Regression Line", lty = 1, col = "green")
summary(lm.fit)
```
The RSE is equal to 4.333
The R-squared value is 0.6918, so 69.18% of the MPG is explained by the regression line.

```{r echo=FALSE}
plot(Auto$weight, Auto$acceleration, xlab = "Weight", ylab = "Acceleration", main = "Acceleration vs Weight")
lm.fit <- lm(Auto$acceleration ~ Auto$weight, Auto)
abline(lm.fit, col = 'red')
legend("topright", legend = "Regression Line", lty = 1, col = "red")
summary(lm.fit)
```

The RSE is equal to 0.1738
The R-squared value is 0.1716, so only 17.16% of the acceleration is explained by regression line for weight.

```{r echo = FALSE}
plot(Auto$mpg, Auto$displacement, xlab = "MPG", ylab = "Displacement", main = "MPG vs Displacement")
lm.fit <- lm(Auto$displacement ~ Auto$mpg, Auto)
abline(lm.fit, col = 'blue')
legend("topright", legend = "Regression Line", lty = 1, col = "blue")
summary(lm.fit)
```
The RSE is equal to 62.14
The R-squared value is 0.6473 so 64.73% of the displacement is explained by the regression line for MPG.

Based on the above plots, there are a number of predictors that can be useful in predicting MPG. 
Almost all of the predictors share a very strong relationship with the MPG predictor, including: horsepower, weight, cylinders, and displacement. 
Acceleration is the only predictor which doesn't have a very strong correlation with MPG. 
This is seen by plotting the correlation matrix along with observing the graphs individually.


